---
title: "Bug Hotels"
author: "Gregory J. Matthews"
date: "2024-04-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data 
Data was collected at 32 total sites with 8 sites each belonging to one of the four treatments (meadow, submerged, typha, tyhyd).  Each individual site was visited twice in 2020 once in late July (July 27-29) and again in early September (September 8-11).  At each site for each visit a sample, we took a sample of insects at each site and the type of insects was recorded.  Each of these insects was grouped into one of5 difference categories (Collector, Shredder, Piercer, Scraper, Predator).  We are priomarily interested in differences in the proportions of these categories at the four different types of sites (meadow, submerged, typha, tyhyd).  Two measurements were dropped due to innacurate oxygen level readings (DO_mg_L 22.7 and 22.3).  Both of these reading were taken on Septemeber 10, 2020 aobut 45 minutes apart and both dropped readings were from the treatment meadow.   Thus all treatments and time points have 8 measurements each except for meadow at the second time point which has only 6 measurements.  This leaves us with a total of 62 observations that were ultiamtely used for analysis.  

```{r echo = FALSE, message = FALSE}
library(rjags)
library(ggmosaic)
library(tidyverse)
cong <- read.csv("/Users/gregorymatthews/Dropbox/bug_hotels/data/cong_master.csv")
library(xtable)
cong %>% group_by(Week, Date) %>% summarize(n = n())
```


# Methods
A Bayesian multinomial mixed effects model was fit to the data with the 

Let $Y_{ijt}$ be the $i$-th observation with $i = 1, cdots, 32$ of the count of the insects/big/whatever the fuck in category $j = 1, \cdots, J$ at time $t = 1, \cdots, T$ where $J =5$ and $T = 2$.  

$$
Y_{i.t} \sim Multinomial(n_{it},p_{it})
$$

$$
p_{ijt} = \frac{\theta_{ijt}}{1 + \sum_{j = 1}^{J-1} \theta_{ijt}}
$$

$$
p_{iJt} = \frac{1}{1 + \sum_{j = 1}^{J-1} \theta_{ijt}}
$$
$$
log(\theta_{ijt}) =  X_{it}\beta_{j} + b_{ij}
$$
for $j = 1, \cdots, J-1$.  Where $X_{it}$ is a $p$ dimensional vector  containing the covariates included in the model (with leading term 1 to account for the intercept) of the for site $i$ at time $t$, $beta_j$ is a $p$ dimensional vectors (including the intercept term), and $b_{ij}$ is a random intercept for category $j$ at site $i$. Covariates include in the model were indicators for treatment (with submerged chosen as the baseline level), Water temperature, Veg_H, and an indicator for week (with week 1 chosen as baseline).  





X <- cong %>% select(I_meadow, I_tyhyd, I_typha, DO_mg_L, Water_Temp_C,Veg_H, I_week)



model{

for (i in 1:N){
  Y[i,] ~ dmulti(p[i,], n[i])

for (j in 1:(J-1)){
  p[i,j] <- theta[i,j]/(1+sum(theta[i,]))
}
  p[i,J] <- 1/(1+sum(theta[i,]))

  for (j in 1:(J-1)){
  theta[i,j] <- exp(inprod(X[i,], beta[,j]) + inprod(Z[i,], b[,j]))
  }
 
 
 
}
 
 
for (d in 1:Dx) {
  for (j in 1:(J-1)){
    beta[d, j] ~ dnorm(0, 0.001)
  }
}

for (d in 1:Dz) {
  for (j in 1:(J-1)){
    b[d, j] ~ dnorm(0, tau2b)
  }
}

tau2b <- 1/sigma2b
sigma2b ~  dnorm(0, 0.001)T(0,10000)
 
 

# Results 


```{r echo = FALSE, message = FALSE}
library(rjags)
library(ggmosaic)
library(tidyverse)
cong <- read.csv("/Users/gregorymatthews/Dropbox/bug_hotels/data/cong_master.csv")
cong %>% group_by(Week, Date) %>% summarize(n = n())
cong %>% group_by(Plot) %>% summarize(n = n()) 

cong %>% filter(DO_mg_L > 20)
#Remove the DO outliers
cong <- cong %>% filter(DO_mg_L <= 20)
#Model variable selection.  Must be in model: Tot_veg_pct and Water_depth_cm
#Add Tot_veg_pct and Water_depth_cm and Water_Temp_C and Veg_H (richness and evenness.  This is just entropy)
cong <- cong %>% select(Date, Site, Plot, Week, Treatment, Lat, Long,
                        Inv_abun, Collector, Shredder, Piercer, Scraper, Predator,
                        DO_mg_L,Tot_veg_pct, Water_depth_cm, Water_Temp_C, Veg_H)

cong <- cong %>% mutate(n_col = Inv_abun*Collector,
                        n_shr = Inv_abun*Shredder,
                        n_pie = Inv_abun*Piercer,
                        n_scr = Inv_abun*Scraper,
                        n_pre = Inv_abun*Predator)


cong$Treatment <- as.factor(cong$Treatment)

#Checking
tot <- cong %>% group_by(Treatment) %>% summarize(sum_n_col = sum(n_col),
                                                  sum_n_shr = sum(n_shr),
                                                  sum_n_pie = sum(n_pie),
                                                  sum_n_scr = sum(n_scr),
                                                  sum_n_pre = sum(n_pre)) 
tot[,-1] <- t(apply(tot[,-1],1,function(x){x/sum(x)}))
```

```{r}
rownames <- tot$Treatment
tab <-tot[,-1] %>% as.matrix() %>% as.table()
row.names(tab) <- c(rownames)
#Rows and columns are not independent
chisq.test(tab)
```

# Some EDA stuff 
```{r}
#Data viz
ggdat <- cong %>% select(Treatment, n_col, n_shr, n_pie, n_scr, n_pre) %>% pivot_longer(names_to = "Category", cols = c("n_col","n_shr","n_pie","n_scr","n_pre"))
ggplot(aes(fill = Category, x = Treatment, y = value), data = ggdat) + geom_bar(position="fill", stat="identity")
```


```{r}
#EDA
names(cong)
table(cong$Date)
table(cong$Site) #All Munuscong
table(cong$Plot)
table(cong$Week)
table(cong$Week, cong$Plot) #Why is MEAD 4 and 5 not repeated?  
table(cong$Treatment)
summary(cong$Inv_abun)
summary(cong$Collector)

cong %>% group_by(Treatment) %>% summarize()

```




```{r echo = FALSE}
#Bayes stuff
cong  <- cong %>% mutate(I_meadow = ifelse(Treatment == "meadow",1,0),
                         I_submerged = ifelse(Treatment == "submerged",1,0),
                         I_tyhyd = ifelse(Treatment == "tyhyd",1,0),
                         I_typha = ifelse(Treatment == "typha",1,0),
                         I_week = ifelse(Week == "W1",0,1))
Y <- cong %>% select(n_col, n_shr, n_pie, n_scr, n_pre) 
Y <- as.matrix(Y)
Y <- matrix(as.integer(Y),nrow = nrow(Y), ncol = ncol(Y))
#Creating the design matrix
#Submerged is baseline
X <- cong %>% select(I_meadow, I_tyhyd, I_typha, DO_mg_L, Water_Temp_C,Veg_H, I_week)
#X <- cong %>% select(I_meadow,I_submerged, I_tyhyd, I_typha,DO_mg_L)
X <- data.frame(int = 1,as.matrix(X))


Z <- matrix(NA, nrow = nrow(cong), ncol = length(unique(cong$Plot)))
for (i in 1:nrow(cong)) {
  for (j in 1:length(unique(cong$Plot))) {
    Z[i, j] <- ifelse(sort(unique(cong$Plot))[j] == cong$Plot[i], 1, 0)
  }
}  
n <- apply(Y, 1, sum)
```

```{r eval = FALSE}
library(rjags)
####################################
#Fixed and random effects
####################################
data <- list(
  N = nrow(Y), # Number of observations
  J = ncol(Y),        # Number of classes
  Dx = ncol(X),  #Number of columns of X
  Dz = ncol(Z), #Number of columns of Z
  Y = Y,     # Response variable
  X = X,      #Predictor matrix Fixed effects
  Z = Z,     #predictor matrix Random effects
  n = n     #number of total bugs at each site
)

model_string <- "
model{

for (i in 1:N){
  Y[i,] ~ dmulti(p[i,], n[i])

for (j in 1:(J-1)){
  p[i,j] <- theta[i,j]/(1+sum(theta[i,]))
}
  p[i,J] <- 1/(1+sum(theta[i,]))

  for (j in 1:(J-1)){
  theta[i,j] <- exp(inprod(X[i,], beta[,j]) + inprod(Z[i,], b[,j]))
  }
 
 
 
}
 
 
for (d in 1:Dx) {
  for (j in 1:(J-1)){
    beta[d, j] ~ dnorm(0, 0.001)
  }
}

for (d in 1:Dz) {
  for (j in 1:(J-1)){
    b[d, j] ~ dnorm(0, tau2b)
  }
}

tau2b <- 1/sigma2b
sigma2b ~  dnorm(0, 0.001)T(0,10000)
 
 


}"
```

```{r eval = FALSE}
setwd("/Users/gregorymatthews/Dropbox/CDSC/bug_hotels/")
write(model_string,"bug_hotels.bug")

# Run JAGS
jags <- jags.model("/Users/gregorymatthews/Dropbox/CDSC/bug_hotels/bug_hotels.bug",
                   data = data,
                   n.chains = 3, 
                   n.adapt = 100
)

update(jags, 1000)

ztest <- jags.samples(jags,c('beta','p','b',"sigma2b"),1000, thin = 1)
z <- jags.samples(jags,c('beta','p','b',"sigma2b"),200000, thin = 200)

#save(z, file = "/Users/gregorymatthews/Dropbox/bug_hotels/z.RData")

```

```{r echo = FALSE}
load("/Users/gregorymatthews/Dropbox/bug_hotels/z_20240417.RData")
```

# Posterior estimates for $\beta$
```{r echo = FALSE}
datbeta <- data.frame(treatment = rep(c("meadow","typhd","typha"), each = 4*3000), 
                   category = rep(c("p_col", "p_shr", "p_pie", "p_scr"),each = 3000),
    beta = c(c(z$beta[2,1,,1],z$beta[2,1,,2],z$beta[2,1,,3]),
           c(z$beta[2,2,,1],z$beta[2,2,,2],z$beta[2,2,,3]),
           c(z$beta[2,3,,1],z$beta[2,3,,2],z$beta[2,3,,3]),
           c(z$beta[2,4,,1],z$beta[2,4,,2],z$beta[2,4,,3]),
           c(z$beta[3,1,,1],z$beta[3,1,,2],z$beta[3,1,,3]),
           c(z$beta[3,2,,1],z$beta[3,2,,2],z$beta[3,2,,3]),
           c(z$beta[3,3,,1],z$beta[3,3,,2],z$beta[3,3,,3]),
           c(z$beta[3,4,,1],z$beta[3,4,,2],z$beta[3,4,,3]),
           c(z$beta[4,1,,1],z$beta[4,1,,2],z$beta[4,1,,3]),
           c(z$beta[4,2,,1],z$beta[4,2,,2],z$beta[4,2,,3]),
           c(z$beta[4,3,,1],z$beta[4,3,,2],z$beta[4,3,,3]),
           c(z$beta[4,4,,1],z$beta[4,4,,2],z$beta[4,4,,3]))
           
    )


ggplot(aes(x = beta), data = datbeta) + geom_density() + facet_grid(treatment ~ category) + geom_vline(xintercept = 0) + ggtitle("All relative to Submerged")
  

```

# Posterior probablities for $\beta > 0$

```{r}
summ <- datbeta %>% group_by(treatment, category) %>% summarize(mean(beta > 0))
summ
```

# Posterior estimates for p
```{r echo = FALSE}
#Plot with 4 rows for each Treatment.  
#And then posterior density plots for each of the categories.
#Using the average level of the random effect, Average Do_mg_L, and Week 0.  
#Find p vectors with average levels 
newX <- X[1:4,]
newX[1:4,2:4] <- 0
diag(newX[2:4,2:4]) <- 1
newX$DO_mg_L <- 1
newX$Water_Temp_C <- mean(X$Water_Temp_C)
newX$Veg_H <- mean(X$Veg_H)


#4 treatments, 5 categories 
results1 <- results2 <- results3 <- list()
for (i in 1:1000){
  results1[[i]] <-
    data.frame(category = rep(c("p_col", "p_shr", "p_pie", "p_scr", "p_pre"),4),                treatment = rep(c("submerged","meadow","typhd","typha"),each = 5),
               p = c(apply(cbind(exp(as.matrix(newX) %*% z$beta[,,i,1]),1),1,function(x) {x / sum(x)})))

results2[[i]] <- data.frame(category = rep(c("p_col", "p_shr", "p_pie", "p_scr", "p_pre"),4),                treatment = rep(c("submerged","meadow","typhd","typha"),each = 5),
               p = c(apply(cbind(exp(as.matrix(newX) %*% z$beta[,,i,1]),2),1,function(x) {x / sum(x)})))

results3[[i]] <- data.frame(category = rep(c("p_col", "p_shr", "p_pie", "p_scr", "p_pre"),4),                treatment = rep(c("submerged","meadow","typhd","typha"),each = 5),
               p = c(apply(cbind(exp(as.matrix(newX) %*% z$beta[,,i,3]),1),1,function(x) {x / sum(x)})))
}

results <- rbind(do.call(rbind,results1),do.call(rbind,results2),do.call(rbind,results3))

results$i <- rep(1:1000,each= 20)
results$chain <- rep(1:3, each = 1000)


ggplot(aes(x = p, y = after_stat(scaled), color = category), data = results) + geom_density() + facet_wrap(~treatment ) +ggtitle("oxygen = 1")
```

```{r echo = FALSE}
#Plot with 4 rows for each Treatment.  
#And then posterior density plots for each of the categories.
#Using the average level of the random effect, Average Do_mg_L, and Week 0.  
#Find p vectors with average levels 
newX <- X[1:4,]
newX[1:4,2:4] <- 0
diag(newX[2:4,2:4]) <- 1
newX$DO_mg_L <- mean(X$DO_mg_L)
newX$Water_Temp_C <- mean(X$Water_Temp_C)
newX$Veg_H <- mean(X$Veg_H)


#4 treatments, 5 categories 
results1 <- results2 <- results3 <- list()
for (i in 1:1000){
  results1[[i]] <-
    data.frame(category = rep(c("p_col", "p_shr", "p_pie", "p_scr", "p_pre"),4),                treatment = rep(c("submerged","meadow","typhd","typha"),each = 5),
               p = c(apply(cbind(exp(as.matrix(newX) %*% z$beta[,,i,1]),1),1,function(x) {x / sum(x)})))

results2[[i]] <- data.frame(category = rep(c("p_col", "p_shr", "p_pie", "p_scr", "p_pre"),4),                treatment = rep(c("submerged","meadow","typhd","typha"),each = 5),
               p = c(apply(cbind(exp(as.matrix(newX) %*% z$beta[,,i,1]),2),1,function(x) {x / sum(x)})))

results3[[i]] <- data.frame(category = rep(c("p_col", "p_shr", "p_pie", "p_scr", "p_pre"),4),                treatment = rep(c("submerged","meadow","typhd","typha"),each = 5),
               p = c(apply(cbind(exp(as.matrix(newX) %*% z$beta[,,i,3]),1),1,function(x) {x / sum(x)})))
}

results <- rbind(do.call(rbind,results1),do.call(rbind,results2),do.call(rbind,results3))

results$i <- rep(1:1000,each= 20)
results$chain <- rep(1:3, each = 1000)


ggplot(aes(x = p, y = after_stat(scaled), color = category), data = results) + geom_density() + facet_wrap(~treatment ) + ggtitle("average values of covariates")
```



```{r}
trt <- c("meadow","submerged","typha","typhd")
cat <- c("p_col","p_pie","p_pre","p_scr","p_shr")

pdiffs <- data.frame()
for (t in 1:(length(trt)-1)){
  for (v in (t+1):length(trt)){
    for (c in 1:length(cat)){
    temp <- results %>% filter(treatment %in% trt[c(t,v)] & category == cat[c]) %>% select(i, chain, treatment, p) %>% pivot_wider(names_from = treatment, values_from = p) 
    pdiffs <- rbind(pdiffs, data.frame(trt1 = trt[t], trt2 = trt[v], cat = cat[c],prob = mean(temp[,3] > temp[,4])))
    }
  }
}

pdiffs %>% pivot_wider(names_from = "cat",values_from = "prob")
```

```{r}
auc <- data.frame()
for (t in 1:(length(trt)-1)){
  for (v in (t+1):length(trt)){
    temp <- results %>% filter(treatment %in% trt[c(t,v)]) %>% select(i, chain, treatment, p, category) %>% pivot_wider(names_from = category, values_from = p) 
    model <- glm(as.factor(treatment) ~ p_col + p_shr + p_pie + p_scr, data = temp, family = "binomial")
    
    predictions <- predict(model, type = "response")
library(pROC)
auc <- rbind(auc, data.frame(trt1 = trt[t], trt2 = trt[v],
  auc = roc(as.factor(temp$treatment), predictions)$auc))
  }
}
auc
```

```{r}
entropy <- results %>% group_by(i,chain, treatment) %>% summarize(entropy = sum(-p*log(p)))
ggplot(aes(x = entropy), data = entropy) + facet_wrap(~treatment) + geom_density()
```

```{r}
entropy <- results %>% group_by(i,chain, treatment) %>% summarize(entropy = sum(-p*log(p)))
ggplot(aes(x = entropy, color = treatment), data = entropy)  + geom_density()
```

<!-- # ```{r} -->
<!-- # test <- results %>% group_by(i,chain, treatment) %>% pivot_wider(names_from = "treatment", values_from = p) %>% group_by(i,chain) %>% mutate(sub_vs_mead = log(submerged) - log(meadow))  -->
<!-- #  -->
<!-- # test -->
<!-- # ``` -->

# Posterior probabilities for p
```{r}
d <- data.frame()
for (cat in c("p_col", "p_shr", "p_pie", "p_scr","p_pre")){
  for (trt in c("meadow","typha","typhd")){
prob <- mean((results %>% filter(category == cat & treatment == "submerged")  %>% select(p))  > (results %>% filter(category == cat & treatment == trt) %>% select(p)))
d <- rbind(d, data.frame(category = cat, 
                         treatment = trt, 
                         prob = prob))
  }}

d
```

# Posterior means and 95% credible intervals for p
```{r}
#Posterior means with 95% credible intervals
results %>% 
  group_by(category, treatment) %>% 
  summarize(mean = mean(p), lower = quantile(p, 0.025), upper = quantile(p, 0.975))

#Posterior means
test <- results %>% 
  group_by(category, treatment) %>% 
  summarize(mean = mean(p)) %>% pivot_wider(names_from = category, values_from = mean) 
```

# Varibility between different plots.  
```{r}
ggdat <- data.frame(sigma2b = c(z$sigma2b[,,1],z$sigma2b[,,2],z$sigma2b[,,3]))
ggplot(aes(x = sigma2b), data = ggdat) + geom_density() 
```

# Appendix

```{r}
#Check convergence. 
conv <- data.frame(iter = rep(1:1000, 3), chain = rep(1:3,each = 1000), sigma2b = c(z$sigma2b[1,,1],z$sigma2b[1,,2],z$sigma2b[1,,3]))
ggplot(aes(x = iter, y = sigma2b, color = factor(chain)),data = conv) + geom_path()

for (i in 1:10){
  for (j in 1:5){
conv <- data.frame(iter = rep(1:1000, 3), chain = rep(1:3,each = 1000), p = c(z$p[i,j,,1],z$p[i,j,,2],z$p[i,j,,3]))
g <- ggplot(aes(x = iter, y = p, color = factor(chain)),data = conv) + geom_path() + ggtitle(paste("i = ",i," j = ",j))
print(g)
}}


#Check ACF plots.
acf(z$sigma2b[1,,1])

for (i in 1:10){
  for (j in 1:5){
acf(z$p[i,j,,1])
    }}

```